Flavour Symmetry Embedded - GLoBES (FaSE-GLoBES)

### 2020-05-18 doc/paper.tex
update docs

\begin{figure}[!h]%
\centering
\includegraphics[width=4.5in]{Figs/FASE-chart_1_2.pdf}
\caption{A scheme to correlate the model parameters with standard neutrino oscillation parameters. The error propagation is implemented in the simulation code up to the spectra analysis.}%
\label{fig:FASE}
\end{figure}

@@

The user can use \textbf{FaSE-GLoBES} to constrain model parameters. The expression $\chi^2$ is used as the default \textbf{GLoBES} setting. \textcolor{red}{[JT: What is the meaning of this $\chi^2$ here?]} In more detail, the $\chi^2$ function is constructed based on a log-likelihood ratio,
%\begin{equation}\label{eq:chi-squared}
\begin{align}\label{eq:chi-squared}
\chi^2(\vec{\theta}^{hyp},\xi_s,\xi_b)=&2\sum_i\left(\eta_i(\vec{\theta}^{hyp},\xi_s,\xi_b)-n_i+n_i\ln\frac{n_i}{\eta_i(\vec{\theta}^{hyp},\xi_s,\xi_b)} \right)\nonumber\\
&+p(\xi_s,\sigma_s)+p(\xi_b,\sigma_b)+\chi^2_{prior},
\end{align}
%\end{equation}
where $i$ runs over the number of bins, $\eta_i(\vec{\theta},\xi_s,\xi_b)$ is the assumed event rate in the $i$th bin and $E_i$ is the central value in this energy bin. The vector $\vec{\theta}$ consists of model or oscillation parameters. The parameters $\xi_s$ and $\xi_b$ are introduced to account for the systematic uncertainties in the normalisation for the signal (subscript $_s$) and background (subscript $_b$) components of the event rate, and are allowed to vary in the fit as nuisance parameters. For a given set of parameters $\vec{\theta}$, the event rate in the $i$th energy bin is calculated as\\
\begin{equation}
\eta_i(\vec{\theta}^{hyp},\xi_s,\xi_b)=(1+\xi_s)\times s_i+(1+\xi_b)\times b_i,
\end{equation}
where $s_i$ and $b_i$ are the expected number of signal and background events in $i$th energy bin, respectively. The nuisance parameters are constrained by the Gaussian prior $p(\xi,\sigma)=\xi^2/\sigma^2$ with corresponding uncertainties $\sigma_s$ and $\sigma_b$ for the signal and background, respectively. Finally, $\chi^2_{prior}$ is a set of Gaussian priors for hypothesis, and is expressed as Eq.~\ref{eq:prior}. After doing all minimisations, the user obtain the $\chi^2$ value for a specific hypothesis $\vec{\theta}^{hyp}$, $\chi^2(\vec{\theta}^{hyp})$.

@@

{\color{red}To show the generality of \textbf{FaSE-GLoBES}, we present the similar result with another model -- \textit{the warped flavor symmetry} (WFS) [Ref???]. This model predict further simplified correlations that the standard oscillation parameters including mixing angles and the CP phase are functions of only two model parameters $\theta_\nu$ and $\phi_\nu$,
\begin{align}
\sin^2\theta_{12}&=\frac{1}{2-\sin2\theta_{\nu}\cos\phi_\nu},\nonumber\\
\sin^2\theta_{13}&=\frac{1}{3}(1+\sin2\theta_\nu\cos\phi_\nu),\nonumber\\
\sin^2\theta_{23}&=\frac{1-\sin2\theta_\nu\sin(\pi/6-\phi_\nu)}{2-\sin2\theta_\nu\cos\phi_\nu},\nonumber\\
J_{CP}&=-\frac{1}{6\sqrt{3}}\cos2\theta_\nu.
\end{align}
The constraint of $\theta_\nu$ and $\phi_\nu$ is presented in Fig~\ref{fig:theta_phi}, in which we use the best fit of NUFIT 4.0 result as the true values ($(\theta_{12},~\theta_{13},~\theta_{23},~\delta,~\Delta m_{21}^2,~\Delta m_{31}^2)=(33.82^\circ,~8.61^\circ,~49.6^\circ,~215^\circ,~7.39\times10^{-5}~\text{eV}^2,~2.525\times10^{-3}~\text{eV}^2)$).}

@@

\begin{figure}[!h]
 \centering
\includegraphics[width=0.5\textwidth]{Figs/SR_th23_dCP.pdf}
\caption{\label{fig:th23_delta}The 2-D exclusion contour for tri-direct littlest seesaw model on the plane of two true standard parameters $\theta_{23}$ and $\delta$, from $1\sigma$ to $5\sigma$. The range for each parameter is taken according to the $3\sigma$ uncertainty in NuFit4.0 results. The black dot denotes the best fit of NuFit4.0 results ($(\theta_{12},~\theta_{13},~\theta_{23},~\delta,~\Delta m_{21}^2,~\Delta m_{31}^2)=(33.82^\circ,~8.61^\circ,~49.6^\circ,~215^\circ,~7.39\times10^{-5}~\text{eV}^2,~2.525\times10^{-3}~\text{eV}^2)$), while the star is the prediction by the tri-direct littlest seesaw model with NuFit4.0 results ($(\theta_{12},~\theta_{13},~\theta_{23},~\delta,~\Delta m_{21}^2,~\Delta m_{31}^2)\sim(36.25^\circ,~8.63^\circ,~47^\circ,~279^\circ,~7.39\times10^{-5}~\text{eV}^2,~2.525\times10^{-3}~\text{eV}^2)$).}
\end{figure}

We can also study on whether a neutrino mass model is excluded or not, assuming different true values for oscillation parameters. In this example, we present testing the TDLS model in various true values of $\theta_{23}$ and $\delta$. We compute the minimal $\chi^2$ value for the model with all four model parameters, varied with the prior Eq.~(\ref{eq:prior}) according to NUFIT4.0 results. In addition, the studied statistics function is exactly given by Eq.~(\ref{eq:chi-squared}), but the true event rate $n_i$ is predicted by a set of oscillation parameters, which will be varied in the code.

As mentioned in Sec.~\ref{sec:chi-squared}, the user needs to consider Wilk's theorem with the different of degrees of freedom between the true theory and the hypothesis, \texttt{dof}. \textcolor{red}{[JT: what is the meaning of the true theory?]} Here the number \texttt{dof} is $2$. And that is our output: \texttt{gsl\_cdf\_chisq\_Qinv(gsl\_cdf\_chisq\_Q(fabs(res),dof),1)}. \textcolor{red}{[I don't understand this sentence.]} We show the result in Fig.~\ref{fig:th23_delta}. More details are presented in the user manual. The code for this example is provided as 
\textbf{th23\_delta.c} in the \textbf{FaSE} distribution.

>>>>

\begin{figure}[!h]%
\centering
\includegraphics[width=7.5in]{Figs/FASE-chart_1_2.pdf}
\caption{A scheme to correlate the model parameters with standard neutrino oscillation parameters. The error propagation is implemented in the simulation code up to the spectra analysis.}%
\label{fig:FASE}
\end{figure}

@@

The user can use \textbf{FaSE-GLoBES} to constrain model parameters. 
{\color{blue}Suppose we have the measurement $\vec{x}$ and the likelihood function $L(\vec{\theta})=P(\vec{x}|\vec{\theta})$ for a set of parameters $\vec{\theta}=(\theta_1,...,\theta_N)$, where $P(\vec{x}|\vec{\theta})$ is the probability function for data $\vec{x}$ in favour of the hypothesis $\vec{\theta}$. The constraint of model parameters can be obtained with the statistis parameter $\chi^2\equiv \ln L(\vec{\theta})$.}
The expression $\chi^2$ is used as the default \textbf{GLoBES} setting. \textcolor{red}{[JT: What is the meaning of this $\chi^2$ here?]} In more detail, the $\chi^2$ function{\color{blue}, following the Poisson distribution,} is constructed based on a log-likelihood ratio,
%\begin{equation}\label{eq:chi-squared}
\begin{align}\label{eq:chi-squared}
\chi^2(\vec{\theta}^{hyp},\xi_s,\xi_b)=&2\sum_i\left(\eta_i(\vec{\theta}^{hyp},\xi_s,\xi_b)-n_i+n_i\ln\frac{n_i}{\eta_i(\vec{\theta}^{hyp},\xi_s,\xi_b)} \right)\nonumber\\
&+p(\xi_s,\sigma_s)+p(\xi_b,\sigma_b)+\chi^2_{prior},
\end{align}
%\end{equation}
where $i$ runs over the number of bins, $\eta_i(\vec{\theta},\xi_s,\xi_b)$ is the assumed event rate in the $i$th bin and $E_i$ is the central value in this energy bin. The vector $\vec{\theta}$ consists of model or oscillation parameters. The parameters $\xi_s$ and $\xi_b$ are introduced to account for the systematic uncertainties in the normalisation for the signal (subscript $_s$) and background (subscript $_b$) components of the event rate, and are allowed to vary in the fit as nuisance parameters. For a given set of parameters $\vec{\theta}$, the event rate in the $i$th energy bin is calculated as\\
\begin{equation}
\eta_i(\vec{\theta}^{hyp},\xi_s,\xi_b)=(1+\xi_s)\times s_i+(1+\xi_b)\times b_i,
\end{equation}
where $s_i$ and $b_i$ are the expected number of signal and background events in $i$th energy bin, respectively. The nuisance parameters are constrained by the Gaussian prior $p(\xi,\sigma)=\xi^2/\sigma^2$ with corresponding uncertainties $\sigma_s$ and $\sigma_b$ for the signal and background, respectively. Finally, $\chi^2_{prior}$ is a set of Gaussian priors for hypothesis, and is expressed as Eq.~\ref{eq:prior}. After doing all minimisations, the user obtain the $\chi^2$ value for a specific hypothesis $\vec{\theta}^{hyp}$, $\chi^2(\vec{\theta}^{hyp})$.

@@

{\color{red}To show the generality of \textbf{FaSE-GLoBES}, we present the similar result with another model -- \textit{the warped flavor symmetry} (WFS)~\cite{Chen:2015jta}. This model predict further simplified correlations that the standard oscillation parameters including mixing angles and the CP phase are functions of only two model parameters $\theta_\nu$ and $\phi_\nu$,
\begin{align}
\sin^2\theta_{12}&=\frac{1}{2-\sin2\theta_{\nu}\cos\phi_\nu},\nonumber\\
\sin^2\theta_{13}&=\frac{1}{3}(1+\sin2\theta_\nu\cos\phi_\nu),\nonumber\\
\sin^2\theta_{23}&=\frac{1-\sin2\theta_\nu\sin(\pi/6-\phi_\nu)}{2-\sin2\theta_\nu\cos\phi_\nu},\nonumber\\
J_{CP}&=-\frac{1}{6\sqrt{3}}\cos2\theta_\nu.
\end{align}
The constraint of $\theta_\nu$ and $\phi_\nu$ is presented in Fig~\ref{fig:theta_phi}, in which we use the best fit of NUFIT 4.0 result as the true values ($(\theta_{12},~\theta_{13},~\theta_{23},~\delta,~\Delta m_{21}^2,~\Delta m_{31}^2)=(33.82^\circ,~8.61^\circ,~49.6^\circ,~215^\circ,~7.39\times10^{-5}~\text{eV}^2,~2.525\times10^{-3}~\text{eV}^2)$).}

@@

\begin{figure}[!h]
 \centering
\includegraphics[width=0.5\textwidth]{Figs/SR_th23_dCP_DUNE_MOMENT.pdf}
\caption{\label{fig:th23_delta}The 2-D exclusion contour for sub-models of tri-direct littlest seesaw on the plane of two true standard parameters $\theta_{23}$ and $\delta$, from $1\sigma$ to $5\sigma$, for DUNE (dotted) and MOMENT (solid). The range for each parameter is taken according to the $3\sigma$ uncertainty in NuFit4.0 results. The black dot denotes the best fit of NuFit4.0 results ($(\theta_{12},~\theta_{13},~\theta_{23},~\delta,~\Delta m_{21}^2,~\Delta m_{31}^2)=(33.82^\circ,~8.61^\circ,~49.6^\circ,~215^\circ,~7.39\times10^{-5}~\text{eV}^2,~2.525\times10^{-3}~\text{eV}^2)$), while the star is the prediction by the tri-direct littlest seesaw model with NuFit4.0 results ($(\theta_{12},~\theta_{13},~\theta_{23},~\delta,~\Delta m_{21}^2,~\Delta m_{31}^2)\sim(36.25^\circ,~8.63^\circ,~47^\circ,~279^\circ,~7.39\times10^{-5}~\text{eV}^2,~2.525\times10^{-3}~\text{eV}^2)$).}
\end{figure}

We can also study on whether a neutrino mass model is excluded or not, assuming different true values for oscillation parameters. In this example, we present testing sub-models of TDLS in various true values of $\theta_{23}$ and $\delta$. We compute the minimal $\chi^2$ value for the model with $r$ and $m_a$, varied with the prior Eq.~(\ref{eq:prior}) according to NUFIT4.0 results. Two parameters $x$ and $\eta$ are fixed at the best fit. In addition, the studied statistics function is exactly given by Eq.~(\ref{eq:chi-squared}), but the true event rate $n_i$ is predicted by a set of oscillation parameters, which will be varied in the code.

As mentioned in Sec.~\ref{sec:chi-squared}, the user needs to consider Wilk's theorem with the different of degrees of freedom between the {\color{blue}assumed} true theory and the hypothesis, \texttt{dof}. {\color{blue}`The assumed true theory' here means that the oscillation parameters assumed for predicting the true event spectrum.} \textcolor{red}{[JT: what is the meaning of the true theory?]} 
%Here the number \texttt{dof} is $2$. 
%And that is our output: \texttt{gsl\_cdf\_chisq\_Qinv(gsl\_cdf\_chisq\_Q(fabs(res),dof),1)}. \textcolor{red}{[I don't understand this sentence.]} 
We show the result in Fig.~\ref{fig:th23_delta}. More details are presented in the user manual. The code for this example is provided as 
\textbf{th23\_delta.c} in the \textbf{FaSE} distribution.
